The Hyperedge Event Model

### 2018-05-24 paper/ba-HEM/ba-sample.tex
revisions until PPC

\begin{aug}
\author{\fnms{Bomin} \snm{Kim}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{bzk147@psu.edu}},
\author{\fnms{Aaron} \snm{Schein}\thanksref{addr2,t3,m1,m2}\ead[label=e2]{aschein@cs.umass.edu}},
\author{\fnms{Bruce} \snm{Desmarais}\thanksref{addr3,t3,m1,m2}\ead[label=e3]{bdesmarais@psu.edu}},
\and
\author{\fnms{Hanna} \snm{Wallach}\thanksref{addr4,t1,m2}
\ead[label=e4]{hanna@dirichlet.net}
%\ead[label=u1,url]{http://www.foo.com}
}
\runauthor{B. Kim et al.}

@@

\subsection{Hypothetical Edges}\label{subsec: Tie}
For every possible sender--receiver pair $(a,r)_{a \neq r}$, we define the ``receiver intensity"---an approximate inverse logit of the probability that edge $d$ is being sent from the sender $a$ to receiver $r$---as a linear combination of network-based statistics:
\begin{equation}
\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr},
\end{equation}
where $\boldsymbol{b}$ is a $P$--dimensional vector of coefficients and $\boldsymbol{x}_{adr}$ is a set of network features which vary depending on the hypotheses regarding canonical processes relevant to network theory such as popularity, reciprocity, and transitivity. In addition, we include intercept term to account for the average (or baseline) number of receivers. We place a Normal prior $\boldsymbol{b} \sim N(\boldsymbol{\mu}_b, \Sigma_b)$.

@@

\subsection{Hypothetical Timestamps}\label{subsec:Time}
Similarly, we consider ``If $a$ were the sender of edge $d$, when would it be sent?" and define the ``timing rate" for the sender $a$
\begin{equation}
\mu_{ad} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ad}),
\end{equation}
where $\boldsymbol{\eta}$ is a $Q$--dimensional vector of coefficients with a Normal prior $\boldsymbol{\eta} \sim N(\boldsymbol{\mu}_\eta,\Sigma_\eta)$, $\boldsymbol{y}_{ad}$ is a set of time-related covariates, e.g., any feature that could affect timestamps of the edge, and $g(\cdot)$ is the appropriate link function such as identity, log, or inverse.

In modeling ``when," we do not directly model the timestamp $t_d$. Instead, we assume that each sender's ``time increment"---i.e., waiting time to next interaction---is drawn from a specific distribution in the exponential family. We define the time increment from edge $d-1$ to edge $d$ as $\tau_{d}$ (i.e., $\tau_{d}= t_d-t_{d-1}$) and specify the distribution of hypothetical timestamps with sender-specfic mean $\mu_{ad}$. Following the generalized linear model (GLM) framework \citep{nelder1972generalized}, we assume the mean and variance of the $\tau_{ad}$ satistify
\begin{equation}
\begin{aligned}
E(\tau_{ad}) &= \mu_{ad},\\
V(\tau_{ad}) &= V(\mu_{ad}),
\end{aligned}
\end{equation}
where $\tau_{ad}$ here is a positive real number. Possible choices of distribution include exponential, Weibull, gamma, and log-normal\footnote{The log-normal distribution is not exponential family but can be used via modeling of $\log(\tau_d)$.} distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the choice of distribution, we may introduce any additional latent variable (e.g., variance parameter $\sigma_\tau^2$ for log-normal and shape parameter $\theta$ for gamma) to account for the variance in time increments. We use $f_\tau(\cdot; \mu)$ and $F_\tau(\cdot; \mu)$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$.
\subsection{Senders, Receivers, and Timestamps}\label{subsec:Observed}
Finally, we choose the actual sender, receivers, and timestamp by selecting the sender--receiver-set pair with the smallest time increment \citep{snijders1996stochastic}:
\begin{equation}
\begin{aligned}
a_d &= \mbox{argmin}_{a}(\tau_{ad}),\\
\boldsymbol{r}_d &= \boldsymbol{u}_{a_d d},\\
t_d &=t_{d-1} + \tau_{a_d d}.
\end{aligned}
\end{equation}
Therefore, it is a sender-driven process in that the receivers and timestamp of an edge is jointly determined by the sender's urgency to send the edge to chosen receivers. Note that our generative process accounts for tied events such that in case of tied events (i.e., multiple senders generated exactly same time increments), we observe all of the tied events without assigning the orders of tied events. Algorithm \ref{alg:generative} summarizes the entire generative process for directed edges with one sender and one or more receivers, outlined in this section.
	\begin{algorithm}[!t]
		\spacingset{1}
			\SetAlgoLined
		\caption{Generative Process: one sender and one or more receivers}
		\begin{algorithmic}
			\FOR{d=1 to D}
			\FOR{a=1 to $A$}
			\FOR{r=1 to $A$ (r $\neq$ a)}
			\STATE	set $\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr}$
			\ENDFOR
			\STATE	draw $\boldsymbol{u}_{ad}  \sim
			\mbox{MB}_G(\boldsymbol{\lambda}_{ad})$
			\STATE		set $\mu_{ad} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ad})$
			\STATE		draw $\tau_{ad} \sim f_\tau(\mu_{ad}, V(\mu_{ad}))$
			\ENDFOR

@@

@@

where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $D$ events ($M \leq D$). If $M=D$ (i.e., no tied events), equation (\ref{eqn:latenttime2}) reduces to equation (\ref{eqn:latenttime}). Note that when we have the latent variable to quantify the variance in time increments (based on the choice of timestamp distribution in Section \ref{subsec:Time}), we also use equation (\ref{eqn:latenttime}) (or equation (\ref{eqn:latenttime2}) in case of tied events) for the additional Metropolis-Hastings update---e.g., $\Pr(\sigma^2_\tau| \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ for log-normal distribution and $\Pr(\theta| \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ for gamma distribution.

@@

	\STATE set initial values of $\boldsymbol{u}, \boldsymbol{b}, \boldsymbol{\eta}$
	\FOR{o=1 to outer}

@@

					\STATE update $u_{adr}$ using Gibbs update --- equation (7)

@@

		\FOR{n=1 to inner1}
			\STATE update $\boldsymbol{b}$ using Metropolis-Hastings --- equation (8)

@@

	\FOR{n=1 to inner2}
			\STATE update $\boldsymbol{\eta}$ using Metropolis-Hastings --- equation (10)
			\STATE (if needed) update $\sigma_\tau^2$ using Metropolis-Hastings --- equation (10)

@@

	\ENDFOR
	\STATE	Summarize the results with: last chain of $\boldsymbol{b}$, and last chain of $\boldsymbol{\eta}$ (and $\sigma_\tau^2$) 
\end{algorithmic}
\label{alg:MCMC}
\end{algorithm}
\subsection{Getting It Right (GiR) Test} \label{subec:GiR}
  Software development is integral to the objective of applying the HEM to real world data. Code review is a valuable process in any research computing context, and the prevalence of software bugs in statistical software is well documented \citep[e.g., ][]{altman2004numerical,mccullough2009accuracy}.  With highly complex models such as the HEM, there are many ways in which software bugs can be introduced and go unnoticed. As such, we present a joint analysis of the integrity of our generative model, sampling equations, and software implementation.

@@

   In the forward samples, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $D$ number of edges, we save these statistics:
   \begin{itemize}
   	\item[1.] Mean of observed receiver sizes $ \lVert \boldsymbol{r}_{d} \rVert_1 $ across $d=1,\ldots,D$,
   	\item[2.] Mean of time increments $\tau_d$ across $d=1,...,D$,
   	\item[3.] $b_p$ value used to generate the samples $p = 1,...,P$,
   	\item[4.] $\eta_q$ value used to generate the samples $q = 1,...,Q$,
   	\item[5.] $\sigma^2_\tau$ value used to generate the samples
   \end{itemize}

   To keep the computational burden of re-running thousands of rounds of inference manageable, we run the GiR using a relatively small artificial sample, consisting of $D=100$ edges, $A=5$ actors, $P=4$ number of edge covariates, and $Q=3$ number of timestamp covariates per each forward or backward samples, using log-normal distibution for the time distribution $f$. We generated $10^5$ sets of forward and backward samples, and then calculated 1,000 quantiles for each of the statistics. We also calculated t-test and Mann-Whitney test p-values in order to test for differences in the distributions generated in the forward and backward samples. Before we calculated these statistics, we thinned our samples by taking every 9th sample starting at the 10,000th sample for a resulting sample size of 10,000, in order to reduce the autocorrelation in the Markov chains. In each case, if we observe a large p-value, this gives us evidence that the distributions generated under forward and backward sampling have the same locations. We depict the GiR results using probability-probability (PP) plots. To compare two samples with a PP-plot we calculate the empirical quantile in each sample of a set of values observed across the two samples, then plot the sets of quantiles in the two samples against each other. If the two samples are from equivalent distributions, the quantiles should line up on a line with zero $y$-intercept, and unit slope (i.e., a 45-degree line). The GiR test results are depicted in Figure \ref{figure:GiRplot}, which show that we pass the test on every statistic.

@@

   	\caption{Probability-Probability plot for the GiR test statistics.}

@@

 	   %\subsection{Data}\label{subsec:Montgomery}
 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with large portion of emails with hyperedges (16.76\%), all of which are a single sender multiple receiver cases, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012.  For this case study,
 	   we formulate the network statistics $\boldsymbol{x}$ and timestamp statistics $\boldsymbol{y}$ and ground them in illustrative examples. We then report a suite of experiments that test our method’s ability to form the posterior distribution over latent variables. 

@@

 	   In the example of email networks, we form the covariate vector $\boldsymbol{x}_{adr}$ using time-varying network statistics based on the time interval prior to and including $t_{d-1}$. Given a set of interaction data, a primary modeling goal lies in determining which characteristics and behaviors of the senders and receivers are predictive of interaction. This email application specifically give rise to the following question: ``To what extent are nodal, dyadic or triadic network effects, as characterized
 	   by past interaction behaviors, relevant to predicting future events?" We employ the network statistics in \cite{PerryWolfe2012}, along with the intercept term, to address the question. Specifically, our time interval tracks 7 days prior to the last email was sent $l_d = (t_{d-1}-7\mbox{ days}, t_{d-1}]$. For $a \in [A], r \in [A]$, and $d \in [D]$, we define the effects

@@

 	   	\item[2.] outdegree of sender $a$: ${x}_{adr2} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
 	   	\item[3.] indegree of receiver $r$: ${x}_{adr3}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime r} = 1)$;
 	   	\item[4.] send: ${x}_{adr4}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
 	   	\item[5.] receive: ${x}_{adr5}=\mbox{send}(r,a)$;
 	   	\item[6.] 2-send: ${x}_{adr6} = \sum_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(h,r)$;
 	   	\item[7.] 2-receive: ${x}_{adr7}= \sum_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(r,h)$;
 	   	\item[8.] sibling: ${x}_{adr8}=\sum_{h \neq a, r} \mbox{send}(h,a)\mbox{send}(h,r)$;
 	   	\item[9.] cosibling: ${x}_{adr9}=\sum_{h \neq a, r} \mbox{send}(a,h)\mbox{send}(r,h)$;
 	   	\item[10.] hyperedge size of sender $a$: ${x}_{adr10}=\sum_{d^\prime: t_{d^\prime} \in l_d} \sum_{r=1}^A I(a_{d^\prime} = a)I(u_{d^\prime r} = 1)$;
 	   	\item[11.] outdegree$*$hyperedge size: ${x}_{adr11} = {x}_{adr2}\times{x}_{adr10};$

@@

 	   where $I(\cdot)$ is an indicator function, and the last covariate is the interaction term between outdegree and hyperedge size. As with the nodal (2, 3, 10, 11) and dyadic (4, 5) effects, the triadic (6--9) effects are designed so that their coefficient have a straightforward interpretation. The statistics ``outdegree" and ``indegree" measures the gregariousness and popularity effect of the node by counting the number of emails sent from $a$ and received by $r$, respectively, within the last 7 days. Moreover, in order to capture individual tendency of writing multicast emails, we include the statistic ``hyperedge size"---the number of emails sent from $a$ within last 7 days where emails with $k$ number of receivers are counted as $k$ separate emails---as a variant of outdegree statistic, accounting for hyperedges. Dyadic statistics ``send" and ``receive" are defined as above such that these covariates measure the number of emails sent from $a$ to $r$ and $r$ to $a$, respectively, within the last 7 days. In the example of triadic statistics, the covariate ``2-send" counts the pairs of emails involving some node $h$ distinct from $a$ and $r$ such that emails from $a$ to $h$ and $h$ to $r$ are both observed within the last 7 days. Other triadic covariates behave similarly, and their interpretations are also analogous, which are illustrated in Figure \ref{figure:netstats}.

@@

 	   	\includegraphics[width=0.55\textwidth]{img/triad.png}	

@@

 	   For time-related covariates $\boldsymbol{y}_{ad}$ in Section \ref{subsec:Time}, we form the covariate vector $\boldsymbol{y}_{ad}$ using the seven statistics which may possibly effect ``time to send the next document". The covariates can be time-invariant (such as gender or manager status) or time-dependent (such as the network statistics used for $\boldsymbol{x}$), and they could depend on $a$ only (nodal effect), $d$ only (temporal effect), or both $a$ and $d$ (nodal \& temporal effect). Specifically, the timestamp statistics are defined as
 	   \begin{itemize}
 	   	\item[1.] intercept: ${y}_{ad1} =1$;
 	   	\item[2.] outdegree of sender $a$: ${y}_{ad2} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
 	   	\item[3.] indegree of sender $a$: ${y}_{ad3}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime a} = 1)$;
 	   	\item[4.] gender of sender $a$: ${y}_{ad4}=I(a= \mbox{female})$;
 	   	\item[5.] manager status of sender $a$: ${y}_{ad5}=I(a= \mbox{County Manager})$;
 	   	\item[6.] weekend of event $d$: ${y}_{ad6} = I(t_{d-1}= \mbox{weekend})$;
 	   	\item[7.] am/pm of event $d$: ${y}_{ad7}= I(t_{d-1}= \mbox{pm})$.
 	   \end{itemize}
 	   Note that our generative process for timestamps in Section \ref{subsec:Time} is sender-orientd process where the sender deterimes ``when to send the email," thus we incorporate network statistics that depends on $a$ only---outdegree of sender $a$ and indegree of sender $a$ (not $r$ this time).

@@

		\caption {Comparison of predictive performance between log-normal and exponential distributions: boxplots for correct sender posterior probability (\textit{left}), distribution of correct receiver posterior probability in log-scale (\textit{middle}), and boxplots for median absolute percentage error (\textit{right}). The dotted blue lines in the left plot represents the correct sender probability expected by random guess---i.e., $1/A=1/18\approx0.056$.}

@@

	p_{di} = \abs*{\frac{\tau^{obs}_d - \tau^{pred}_{di}}{\tau^{obs}_d}},
	and we present boxplots for the MdAPE estimates on the right plot. For better visualization, we plot MdAPE estimates in log-scale since they are highly right-skewed. As expected, we have huge benefit in the performance of timestamp prediction when we choose log-normal distribution over exponential. This difference can be explained by our variance estimate $\hat{\sigma}_\tau^2= 14.093$ in Section \ref{subsec:Result_email}, because exponential distribution omits the extra information on time increments contained in the large variance estimate. As illustrated above, we can use this out-of-sample prediction task for two usage---1) provide an effective answer to the question ``how are we filling in the missing information of the emails?" and 2) offer one standard way to determine the time increment distribution in Section \ref{subsec:Time}.

>>>>

\begin{aug}
\author{\fnms{Bomin} \snm{Kim}\thanksref{addr1,t1,t2,m1}\ead[label=e1]{bzk147@psu.edu}},
\author{\fnms{Aaron} \snm{Schein}\thanksref{addr2,t3,m1,m2}\ead[label=e2]{aschein@cs.umass.edu}},
\author{\fnms{Bruce} \snm{A. Desmarais}\thanksref{addr3,t3,m1,m2}\ead[label=e3]{bdesmarais@psu.edu}},
\and
\author{\fnms{Hanna} \snm{Wallach}\thanksref{addr4,t1,m2}
\ead[label=e4]{hanna@dirichlet.net}
%\ead[label=u1,url]{http://www.foo.com}
}
\runauthor{B. Kim et al.}

@@

\subsection{Hypothetical Edges}\label{subsec: Tie}
For every possible sender--receiver pair $(a,r)_{a \neq r}$, we define the ``receiver intensity"---an approximate inverse logit of the probability that edge $d$ is being sent from the sender $a$ to receiver $r$---as a linear combination of network-based statistics:
\begin{equation}
\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr},
\end{equation}
where $\boldsymbol{b}$ is a $P$--dimensional vector of coefficients and $\boldsymbol{x}_{adr}$ is a set of receiver selection features which vary depending on the hypotheses regarding canonical processes relevant to network theory such as popularity, reciprocity, and transitivity. In addition, we include intercept term to account for the average (or baseline) number of receivers. We place a Normal prior $\boldsymbol{b} \sim N(\boldsymbol{\mu}_b, \Sigma_b)$.

@@

\subsection{Hypothetical Timestamps}\label{subsec:Time}
Similarly, we consider ``If $a$ were the sender of edge $d$, when would it be sent?" and define the ``timing rate" for the sender $a$
\begin{equation}
\mu_{ad} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ad}),
\end{equation}
where $\boldsymbol{\eta}$ is a $Q$--dimensional vector of coefficients with a Normal prior $\boldsymbol{\eta} \sim N(\boldsymbol{\mu}_\eta,\Sigma_\eta)$, $\boldsymbol{y}_{ad}$ is a set of event timing features, e.g., any covariates that could affect timestamps of the edge, and $g(\cdot)$ is the appropriate link function such as identity, log, or inverse.

In modeling ``when," we do not directly model the timestamp $t_d$. Instead, we assume that each sender's ``time increment"---i.e., waiting time to next interaction---is drawn from a specific distribution in the exponential family. We define the time increment from edge $d-1$ to edge $d$ as $\tau_{d}$ (i.e., $\tau_{d}= t_d-t_{d-1}$) and specify the distribution of hypothetical timestamps with sender-specfic mean $\mu_{ad}$. Following the generalized linear model (GLM) framework \citep{nelder1972generalized}, we assume the mean and variance of the $\tau_{ad}$ satistify
\begin{equation}
\begin{aligned}
E(\tau_{ad}) &= \mu_{ad},\\
V(\tau_{ad}) &= V(\mu_{ad}),
\end{aligned}
\end{equation}
where $\tau_{ad}$ here is a positive real number. Possible choices of distribution include exponential, Weibull, gamma, and log-normal\footnote{The log-normal distribution is not exponential family but can be used via modeling of $\log(\tau_d)$.} distributions, which are commonly used in time-to-event modeling \citep{rao2000applied,rizopoulos2012joint}. Based on the choice of distribution, we may introduce any additional latent variable (e.g., variance parameter $\sigma_\tau^2$ for log-normal and shape parameter $\theta$ for gamma) to account for the variance in time increments. We use $f_\tau(\cdot; \mu)$ and $F_\tau(\cdot; \mu)$ to denote the probability density function (p.d.f) and cumulative density function (c.d.f), respectively, with mean $\mu$.
\subsection{Senders, Receivers, and Timestamps}\label{subsec:Observed}
Finally, we choose the actual sender, receivers, and timestamp by selecting the sender--receiver-set pair with the smallest time increment \citep{snijders1996stochastic}:
\begin{equation}
\begin{aligned}
a_d &= \mbox{argmin}_{a}(\tau_{ad}),\\
\boldsymbol{r}_d &= \boldsymbol{u}_{a_d d},\\
t_d &=t_{d-1} + \tau_{a_d d}.
\end{aligned}
\end{equation}
Therefore, it is a sender-driven process in that the receivers and timestamp of an edge is jointly determined by the sender's urgency to send the edge to chosen receivers. Note that our generative process accounts for tied events such that in case of tied events (i.e., multiple senders generated exactly same time increments), we observe all of the tied events without assigning the orders of tied events. Algorithm \ref{alg:generative} summarizes the entire generative process for directed edges with one sender and one or more receivers, outlined in this section.
	\begin{algorithm}[!t]
		\spacingset{1}
			\SetAlgoLined
		\caption{Generative Process: one sender and one or more receivers}
		\begin{algorithmic}
				\STATE \textbf{Input}: number of edges and nodes $(D, A)$, covariates $(\boldsymbol{x}, \boldsymbol{y})$, and the coefficients $(\boldsymbol{b}, \boldsymbol{\eta})$
				\vskip 0.1in
			\FOR{d=1 to D}
			\FOR{a=1 to $A$}
			\FOR{r=1 to $A$ (r $\neq$ a)}
			\STATE	set $\lambda_{adr} = {\boldsymbol{b}}^{\top}\boldsymbol{x}_{adr}$
			\ENDFOR
			\STATE	draw $\boldsymbol{u}_{ad}  \sim
			\mbox{MB}_G(\boldsymbol{\lambda}_{ad})$
			\STATE		set $\mu_{ad} = g^{-1}(\boldsymbol{\eta}^\top \boldsymbol{y}_{ad})$
			\STATE		draw $\tau_{ad} \sim f_\tau(\mu_{ad}, V(\mu_{ad}))$
			\ENDFOR

@@

\label{eqn:latentedge}

@@

where $t_1^*,\ldots,t_M^*$ are the unique timepoints across $D$ events ($M \leq D$). If $M=D$ (i.e., no tied events), equation (\ref{eqn:latenttime2}) reduces to equation (\ref{eqn:latenttime}). Note that when we have the latent variable to quantify the variance in time increments $V(\cdot)$ (based on the choice of timestamp distribution in Section \ref{subsec:Time}), we also use equation (\ref{eqn:latenttime}) (or equation (\ref{eqn:latenttime2}) in case there exist tied events) for the additional Metropolis-Hastings update---e.g., $\Pr(\sigma^2_\tau| \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ for log-normal distribution and $\Pr(\theta| \boldsymbol{\eta},\boldsymbol{u}, \boldsymbol{y},\boldsymbol{a}, \boldsymbol{r},\boldsymbol{t})$ for gamma distribution.

@@

	\STATE \textbf{Input}: number of outer and inner iterations $(O, I_1, I_2)$ and initial values of $(\boldsymbol{u}, \boldsymbol{b}, \boldsymbol{\eta})$
				\vskip 0.1in
	\FOR{o=1 to O}

@@

					\STATE update $u_{adr}$ using Gibbs update --- equation (\ref{eqn:latentreceiver})

@@

		\FOR{n=1 to $I_1$}
			\STATE update $\boldsymbol{b}$ using Metropolis-Hastings --- equation (\ref{eqn:latentedge})

@@

	\FOR{n=1 to $I_2$}
			\STATE update $\boldsymbol{\eta}$ using Metropolis-Hastings --- equation (\ref{eqn:latenttime}) or equation (\ref{eqn:latenttime2}) 
			\STATE (if needed) update the latent variable for $V(\cdot)$ using Metropolis-Hastings --- equation (\ref{eqn:latenttime}) or equation (\ref{eqn:latenttime2}) 

@@

	\ENDFOR
	\STATE	summarize the results with the last chain of $\boldsymbol{b}$ and $\boldsymbol{\eta}$
\end{algorithmic}
\label{alg:MCMC}
\end{algorithm}
\subsection{Getting It Right (GiR) Test} \label{subec:GiR}
  Software development is integral to the objective of applying the HEM to real world data. Code review is a valuable process in any research computing context, and the prevalence of software bugs in statistical software is well documented \citep[e.g., ][]{altman2004numerical,mccullough2009accuracy}.  With highly complex models such as the HEM, there are many ways in which software bugs can be introduced and go unnoticed. As such, we present a joint analysis of the integrity of our generative model, sampling equations, and software implementation.

@@

   In the forward samples, both observable and unobservable variables are generated using Algorithm \ref{alg:generative}. In the backward samples, unobservable variables are generated using the sampling equations for inference, which we derived in Section \ref{subsec:conditionaldist}. For each forward and backward sample that consists of $D$ number of edges, we save these statistics:
   \begin{itemize}
   	\item[1.] Mean of observed receiver sizes $ \lVert \boldsymbol{r}_{d} \rVert_1 $ across $d=1,\ldots,D$,
   	\item[2.] Mean of time increments $\tau_d$ across $d=1,...,D$,
   	\item[3.] $b_p$ value used to generate the samples $p = 1,...,P$,
   	\item[4.] $\eta_q$ value used to generate the samples $q = 1,...,Q$,
   	\item[5.] $\sigma^2_\tau$ value used to generate the samples in log-normal distribution
   \end{itemize}

   To keep the computational burden of re-running thousands of rounds of inference manageable, we run the GiR using a relatively small artificial sample, consisting of $D=100$ edges, $A=5$ actors, $P=4$ number of edge covariates, and $Q=3$ number of timestamp covariates per each forward or backward samples, using log-normal distibution for the time increments $f_\tau$. We generated $10^5$ sets of forward and backward samples, and then calculated 1,000 quantiles for each of the statistics. We also calculated t-test and Mann-Whitney test p-values in order to test for differences in the distributions generated in the forward and backward samples. Before we calculated these statistics, we thinned our samples by taking every 9th sample starting at the 10,000th sample for a resulting sample size of 10,000, in order to reduce the autocorrelation in the Markov chains. In each case, if we observe a large p-value, this gives us evidence that the distributions generated under forward and backward sampling have the same locations. We depict the GiR results using probability-probability (PP) plots. To compare two samples with PP plots we calculate the empirical quantile in each sample of a set of values observed across the two samples, then plot the sets of quantiles in the two samples against each other. If the two samples are from equivalent distributions, the quantiles should line up on a line with zero $y$-intercept, and unit slope (i.e., a 45-degree line). The GiR test results are depicted in Figure \ref{figure:GiRplot}, which show that we pass the test on every statistic.

@@

   	\caption{Probability-probability (PP) plots for the GiR test statistics.}

@@

 	   Our data come from the North Carolina county government email dataset collected by \cite{ben2017transparency} that includes internal email corpora covering the inboxes and outboxes of managerial-level employees of North Carolina county governments. Out of over twenty counties, we chose Montgomery County to 1) test our model using data with large proportion of hyperedges (16.76\%), all of which are emails sent from one sender to two or more receivers, and 2) limit the scope of this initial application. To summarize, Montgomery County email network contains 680 emails, sent and received by 18 department managers over a period of 3 months (March--May) in 2012. For this case study,
 	   we formulate the edge covariates $\boldsymbol{x}$ and timestamp covariates $\boldsymbol{y}$ and ground them in illustrative examples. We then report a suite of experiments that test the HEM's ability to form the posterior distribution over latent variables. 

@@

 	   \subsubsection{Edge covariates}
 	   Given a set of network data, a primary modeling goal of the generative process for edges lies in determining which characteristics and behaviors are predictive of the receiver selection. This email application specifically give rise to the following question: ``To what extent are nodal, dyadic or triadic network effects relevant to predicting future emails?" As an illustrative example, we form the receiver selection features $\boldsymbol{x}$ for Montgomery County email data using some nodal statistics and interval-based network statistics. First, we include the gender information of sender and receiver, as well as their homophily indicator. For dyadic and triadic network effects, we summarize past interaction behaviors based on the time interval prior to and including $t_{d-1}$. Specifically, our time interval tracks 7 days prior to the last email was sent $l_d = (t_{d-1}-7\mbox{ days}, t_{d-1}]$. For $a \in [A], r \in [A]$, and $d \in [D]$, we define 14 covariates for $\boldsymbol{x}_{adr}$:

@@

 	   	 \item[2.] gender of sender: ${x}_{adr2} = I(\mbox{gender of sender } a= \mbox{female});$
 	   	 \item[3.] gender of receiver: ${x}_{adr3} = I(\mbox{gender of receiver } r= \mbox{female});$
 	   	 \item[4.] gender homophily: ${x}_{adr4} = I({x}_{adr2}={x}_{adr3});$ 	   		
 	   	\item[5.] outdegree of sender: ${x}_{adr5} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
 	   	\item[6.] indegree of receiver: ${x}_{adr6}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime r} = 1)$;
 	   	\item[7.] hyperedge size of sender: ${x}_{adr7}=\sum_{d^\prime: t_{d^\prime} \in l_d} \sum_{r=1}^A I(a_{d^\prime} = a)\,I(u_{d^\prime r} = 1)$;
 	  \item[8.] interaction between outdegree and hyperedge size: ${x}_{adr8} = {x}_{adr5}\times{x}_{adr7};$
 	   	\item[9.] send: ${x}_{adr9}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)\,I(u_{d^\prime r} = 1)$;
 	   	\item[10.] receive: ${x}_{adr10}=\mbox{send}(r,a)$;
 	   	\item[11.] 2-send: ${x}_{adr11} = \sum_{h \neq a, r} \mbox{send}(a,h)\,\mbox{send}(h,r)$;
 	   	\item[12.] 2-receive: ${x}_{adr12}= \sum_{h \neq a, r} \mbox{send}(h,a)\,\mbox{send}(r,h)$;
 	   	\item[13.] sibling: ${x}_{adr13}=\sum_{h \neq a, r} \mbox{send}(h,a)\,\mbox{send}(h,r)$;
 	   	\item[14.] cosibling: ${x}_{adr14}=\sum_{h \neq a, r} \mbox{send}(a,h)\,\mbox{send}(r,h)$;

@@

 	   where $I(\cdot)$ is an indicator function. The network statistics (5--14) are designed so that their coefficient have a straightforward interpretation. The statistics ``outdegree" and ``indegree" measures the gregariousness and popularity effect of the node by counting the number of emails sent from $a$ and received by $r$, respectively, within the last 7 days. Moreover, in order to capture individual tendency of send emails to two or more receivers, we include the statistic ``hyperedge size"---the number of emails sent from $a$ within last 7 days where emails with $k$ number of receivers are counted as $k$ separate emails---as a variant of outdegree statistic, accounting for hyperedges. We also include the interaction term between outdegree and hyperedge size. Next, we employ dyadic and triadic network statistics from \cite{PerryWolfe2012}. Dyadic statistics``send" and ``receive" are defined as above such that these covariates measure the number of emails sent from $a$ to $r$ and $r$ to $a$, respectively, within the last 7 days. In the example of triadic statistics, the covariate ``2-send" counts the pairs of emails involving some node $h$ distinct from $a$ and $r$ such that emails from $a$ to $h$ and $h$ to $r$ are both observed within the last 7 days. Other triadic covariates behave similarly, and their interpretations are also analogous, which are illustrated in Figure \ref{figure:netstats}.

@@

 	   	\includegraphics[width=0.65\textwidth]{img/triad-1.png}	

@@

 	    	   \subsubsection{Timestamp covariates}
 	   For the event timing features $\boldsymbol{y}$ introduced in Section \ref{subsec:Time}, we identify a set of covariates which may possibly effect ``time to the next email." Similar to the edge covariates, we include nodal statistics which are time-invariant (such as gender or manager status) or time-dependent (such as the network statistics used for $\boldsymbol{x}$). In addition, we select some edge-specific covariates based on the temporal aspect of the $(d-1)^{th}$ email---e.g., whether the previous email was sent (1) during the weekend and (2) before or past midday (AM/PM)---since we expect the email interactions within county government to be less active during the weekend and in the evening. To be specific, the timestamp statistics are defined as
 	   \begin{itemize}
 	   	\item[1.] intercept: ${y}_{ad1} =1$;
 	   \item[2.] gender of sender: ${y}_{ad2}=I(\mbox{gender of sender }a= \mbox{female})$;
 	   \item[3.] manager status of sender: ${y}_{ad3}=I(\mbox{sender }a \mbox{ is the County Manager})$;
 	   	\item[4.] outdegree of sender: ${y}_{ad4} =\sum_{d^\prime: t_{d^\prime} \in l_d} I(a_{d^\prime} = a)$;
 	   	\item[5.] indegree of sender: ${y}_{ad5}=\sum_{d^\prime: t_{d^\prime} \in l_d} I(u_{d^\prime a} = 1)$;
 	   	\item[6.] weekend indicator of previous email: ${y}_{ad6} = I(t_{d-1} \mbox{ is during the } \mbox{weekend})$;
 	   	\item[7.] AM/PM of previous email: ${y}_{ad7}= I(t_{d-1} \mbox{ in } \mbox{PM})$.
 	   \end{itemize}
 	   Note that our generative process for timestamps in Section \ref{subsec:Time} is sender-orientd process where the sender deterimes ``when to send the email," thus we incorporate network statistics that depends on $a$ only---outdegree of sender $a$ and indegree of sender $a$ (not $r$ this time).

@@

		\caption {Comparison of predictive performance between log-normal and exponential distributions: boxplots for correct sender posterior probability (\textit{left}), $F_1$ scores from receiver predictions (\textit{middle}), and boxplots for median absolute percentage error (\textit{right}). The dotted blue lines in the left plot represents the correct sender probability expected by random guess---i.e., $1/A=1/18\approx0.056$.}

@@

	\subsection{Posterior Predictive Checks}\label{subsec:PPC_email} 	   
In this section, we perform posterior predictive checks (PPC) \citep{rubin1984bayesianly} to evaluate the appropriateness of our model specification for Montgomery County email data. We formally generated entirely new data, by simulating $N=500$ synthetic email data $\{(a_{d}, \boldsymbol{r}_{d}, t_{d})\}_{d=1}^D$ from the genenerative process in Section \ref{sec:generative process}, conditional upon a set of inferred latent variables from the inference in Section \ref{subsec:Result_email}, where the pseudocode is outlined in Algorithm \ref{alg:PPC}. For the test of goodness-of-fit in terms of network dynamics, we use multiple statistics that summarize meaningful aspects of the data: outdegree distribution, indegree distribution, receiver size distribution, and probability-probability (PP) plot for time increments. 
